Modeling Liver with Glisson's Capsule in Real-Time Medical Simulations

### 2013-05-30 Doc/Vrst2013/vrst2013.tex
more fixes in the paper
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@578 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\title{Complete Real-Time Liver Model Including Glisson's Capsule, Vascularization and Parenchyma}
%\subtitle{}

@@

\maketitle
\begin{abstract}
The accurate mechanical modeling of liver is of a paramount interest in simulation-based operation planning 
and computer-aided per-operative guidance. However, the realistic simulation of liver behaviour is a challenging task, 
since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, each having different mechanical properties. While the 
mechanical influence of vessels on both global and local behaviour of the organ has been measured and modelled previously~\cite{Peterlik2012}, only local 
effects have been studied in the case of Glisson's capsule~\cite{Hollenstein2006,Ahn2010}.

@@

The mechanical importance of the vascular structures in liver is studied in~\cite{Peterlik2012}. It shows that the 
influence of vessels on the mechanical behaviour of the organ is significant. %, mainly if large deformations occur. 
%Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
%the authors propose a composite model allowing for real-time simulation of entire liver with vascularization.
Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
the authors propose a composite model allowing for real-time simulation.

@@

\subsection{Parenchyma and Vascularization} %{{{

@@

\subsection{Coupling Between Capsule and Parenchyma} %{{{
The literature reports high cohesion between capsule and parenchyma.
Based on this property we assume there is no relative motion of the capsule \wrt\ the parenchyma.
Although an arbitrary surface mesh could be used to model the capsule, we exploit 
the fact that the parenchyma is modeled by tetrahedral elements having
triangular faces. Thus, as the boundary of the volumetric mesh is already
triangulated, we simply employ the triangles on the mesh surface to model the capsule.

@@

A good example of such a local deformation is the aspiration test
where the response of liver exposed locally to a negative pressure is measured.
In~\cite{Hollenstein2006} it is reported that the Glisson's
capsule has a non-negligible influence on the response of the tissue and modeling
only the parenchyma leads to an overestimation of the deformation. 
The aspiration device consists of a tube 1\,cm in diameter and allows
control of internal pressure in the tube. The test is performed by
attaching the tube to the tissue and measuring the tissue response. We
set up a simulation in SOFA to reproduce the experiment (see
Fig.~\ref{fig-aspiration1}: we meshed a 15$\times$15$\times$15\,mm$^3$ 
cube representing the tissue 
resulting in 2648 tetrahedra. Then we attached a 1\,cm tube 
and applied a pressure of 3\,kPa inside the tube. The interaction between the tube and 
tissue was modelled with friction contact.

@@

In the simulations we used material properties reported in the literature: for the 
Glisson's capsule we used Young's modulus $E_c$=8.22\,MPa and thickness $t_c$=20\,$\mu$m~\cite{Umale2011}.
Since the values for parenchyma elasticity quite vary (usually between 2\,kPa to 5\,kPa), we perform the simulation 
with $E_p$=3.5\,kPa.

In Fig.~\ref{fig-aspiration2} the profiles of cuts in the middle of the
test cube are presented showing significantly larger deformation in the model without capsule.
We can observere that the deformation without the capsule is indeed overestimated approximately by the factor of 2. This is in perfect agreement with the results reported in~\cite{Hollenstein2006}

Moreover, if we neglect the time-dependent behaviour we are able to reproduce
the measurements presented in~\cite{Hollenstein2006,Nava2008} with less than 1\% error.
In this case the simulation is performed with the pressure of 30\,kPa and with
the following elastic parameters: $E_p$=30\,kPa for parenchyma and $E_c$=3\,MPa
and $t$=93\,$\mu$m for capsule. All the values lie within the measurements
specified in the paper. \TG{add figure?}

@@

Presently we are working on the validation of the model also in context of
global deformations where we believe the capsule plays also an essential role.
This task is however complicated by the lack of biomechanical literature
dealing with the response of the capsule.
%The influence of the capsule on local level has been published in literature
%before. However, to our knowledge it's influence on global scale deformation
%has not been studied yet. We believe that the capsule has non-negligible impact
%also on global level.

Our goal is to create complete liver model suitable for real-time medical
simulations. For this a model accounting for all three main constituents
(parenchyma, vascularization and capsule) is necessary. To achieve that we have
integrated our model with the with the vasularization method proposed
in~\cite{Peterlik2012}. In our future work we need to validate the model in the
context of global and local deformations. This task is challenging because
comparing in vivo measurements with the simulation is difficult due to complex
boundary conditions. 
\TG{ex-vivo: no information abou vessels; phantoms: cannot model capsule}

@@

In the paper we presented a liver model accounting for the Glisson's capsule, a component that is often overlooked. Each constituent being modelled with a different type 
of finite elements: tetrahedral corotational model used for the parenchyma and
and CST membrane elements used for the capsule. 
The model was validated for local displacements based on the the aspiration test presented in~\cite{Hollenstein2006} 
was successfully reproduced, showing a good agreement between the simulation and experiments.

>>>>

\title{Modeling Liver with Glisson's Capsule in Real-Time Medical Simulations}
%\subtitle{}

@@

\maketitle
\begin{abstract}
  \TODO{}
The accurate mechanical modeling of liver is of a paramount interest in simulation-based operation planning 
and computer-aided per-operative guidance. However, the realistic simulation of liver behaviour is a challenging task, 
since the organ is composed of three constituents: parenchyma, vascularization and Glisson's capsule, each having different mechanical properties. While the 
mechanical influence of vessels on both global and local behaviour of the organ has been measured and modelled previously~\cite{Peterlik2012}, only local 
effects have been studied in the case of Glisson's capsule~\cite{Hollenstein2006,Ahn2010}.

@@

\TG{The following paragraph is optional. Unnecessary but adds to the global context}
The mechanical importance of the vascular structures in liver is studied in~\cite{Peterlik2012}. It shows that the 
influence of vessels on the mechanical behaviour of the organ is significant. %, mainly if large deformations occur. 
%Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
%the authors propose a composite model allowing for real-time simulation of entire liver with vascularization.
Since a detailed modeling of the vessels would be extremely costly (mainly because of the small thickness of the vessel wall, 
the authors propose a composite model allowing for real-time simulation.

@@

\subsection{Parenchyma} %{{{

@@

\subsection{Coupling Between Capsule and\\Parenchyma} %{{{
The literature reports high cohesion between capsule and parenchyma.
Based on this property we assume there is no relative motion of the capsule \wrt\ the parenchyma.
Although an arbitrary surface mesh could be used to model the capsule, we exploit 
the fact that the parenchyma is modeled by tetrahedral elements having
triangular faces. Thus, as the boundary of the volumetric mesh is already
triangulated, we simply employ the triangles on the mesh surface to model the capsule.

@@

A good example of such a local deformation is the aspiration test
where the response of liver exposed locally to a negative pressure is measured.
The aspiration device consists of a tube 1\,cm in diameter and allows
control of internal pressure in the tube. The test is performed by
attaching the tube to the tissue and measuring the tissue response. We
set up a simulation in SOFA to reproduce the experiment (see
Fig.~\ref{fig-aspiration1}: we meshed a 15$\times$15$\times$15\,mm$^3$ 
cube representing the tissue 
resulting in 2648 tetrahedra. Then we attached a 1\,cm tube 
and applied a pressure of 3\,kPa inside the tube. The interaction between the tube and 
tissue was modelled with friction contact.

@@

In the simulations we used material properties for a pig liver reported in the literature: for the 
Glisson's capsule we used Young's modulus $E_c$=8.22\,MPa and thickness $t_c$=20\,$\mu$m~\cite{Umale2011}.
Since the values for parenchyma elasticity quite vary (usually between 2\,kPa to 5\,kPa), we perform the simulation 
with $E_p$=3.5\,kPa.

In~\cite{Hollenstein2006} it is reported that the Glisson's
capsule has a non-negligible influence on the response of the tissue and modeling
only the parenchyma leads to an overestimation of the deformation up to the factor of 3. 
In Fig.~\ref{fig-aspiration2} the profiles of cuts in the middle of the
test cube are presented showing significantly larger deformation in the model without capsule.
We can observere that the deformation without the capsule is indeed overestimated approximately by the factor of 2. This is in good agreement with the results reported in~\cite{Hollenstein2006}

Moreover, if we neglect the time-dependent behaviour we are able to reproduce
the measurements presented in~\cite{Hollenstein2006,Nava2008} with less than 1\% error.
In this case the simulation was performed with the pressure of 30\,kPa and with
the following elastic parameters: $E_p$=30\,kPa for parenchyma and $E_c$=3\,MPa
and $t$=93\,$\mu$m for capsule. All the values lie within the measurements
specified in the paper. \TG{add figure like Fig. 2?}

@@

Presently we are working on the validation of the model also in context of
global deformations where we believe the capsule also plays an essential role.
This task is however complicated by the lack of biomechanical literature
dealing with the response of the capsule.
%The influence of the capsule on local level has been published in literature
%before. However, to our knowledge it's influence on global scale deformation
%has not been studied yet. We believe that the capsule has non-negligible impact
%also on global level.

Our goal is to create complete liver model suitable for real-time medical
simulations. For this a model accounting for all three main constituents
(parenchyma, vascularization and capsule) is necessary. To achieve that we have
integrated our model with the vasularization method proposed
in~\cite{Peterlik2012}. In our future work we need to validate the model in the
context of global and local deformations. This task is challenging because
comparing in vivo measurements with the simulation is difficult due to complex
boundary conditions. 
\TG{ex-vivo: no information abou vessels; phantoms: cannot model capsule}

@@

In the paper we presented a liver model accounting for the Glisson's capsule, a component that is often overlooked. Each constituent being modelled with a different type 
of finite elements: tetrahedral corotational model used for the parenchyma and
and CST membrane elements used for the capsule. 
The model was validated in the context for local displacements based on the the aspiration test presented in~\cite{Hollenstein2006}. The aspiration test
was successfully reproduced, showing a good agreement between the simulation and experiments.
